Playing Atari Breakout with Keras and R

### 2018-07-13 article/article.Rmd
rename funs and add play episode to article

>>>>

## Playing a game

We can now implement a function to play an *episode*, ie. a full match, of the game.
First we implemented an auxiliary function that will reset the environment and return a first valid state.

```{r}
reset_env <- function(env) {
  list(
    s_t = env$reset() %>%
      preprocess() %>%                     # do preprocessing
      abind::abind(., ., ., ., along = 3), # stack the first frame 4 times
    terminal = FALSE
  )
}
```

Now we will implement the `play_episode` function. Most of the logic of the DQN 
algorithm is implemented here. It's inputs are:

* `env`: an environment created by `gym$make`.
* `epsilon`: a numeric value (between 0 and 1) giving the probability of performing a random action. If 1 will perform all actions randomly and if 0, all actions will be chosen by the model.
* `models`: the models (we will detail in the next section)
* `experience`: an instance of the `Experience` class defined in the previous section.
* `train`: a boolean indicating if we will perform (or not) a training step. (we will also detail it later).
* `render`: a boolean indicating if you want to render the game. We recommend disabling it since it may slow down training.

The returned value of `play_episode` is the score attended by the algorithm.

<aside>
It's important to be familiar with the `%<-%` operator from the [zeallot package](https://github.com/r-lib/zeallot). The `%<-%` operator unpacks a list of values while assigning, for example:

```{r}
c(x, y) %<-% list(0, 1)
#> x
#[1] 0
#> y
#[1] 1
```
</aside>

```{r}
play_episode <- function(env, epsilon, models, experience, train, render = FALSE) {

  # resets the envinronment to the initial state
  c(s_t, terminal) %<-% reset_env(env)
  score <- 0 
  # while loop until the game ends - lost all 5 lifes or attended the maximum 
  # number of 10000 frames.
  while(!terminal) {

    # this if-else block will choose if the action performed will be random or
    # if the will use the model to choose the best action. 
    # the epsilon parameter controls the exploration/exploitation trade-off
    if (runif(1) < epsilon) {
      action <- env$action_space$sample()
    } else {
      score_predictions <- predict(models$score_model, abind(s_t, along = 0.1))
      action <- which.max(score_predictions) - 1L
    }

    # performs the selected action on the environment
    c(frame, reward, terminal, lives) %<-% env$step(action)
    if (render) env$render() # render the game

    # stacks last 3 frames from the previous state with the frame 
    # obtained at this timestep.
    s_t1 <- abind::abind(preprocess(frame), s_t[,,1:3], along = 3)

    # stores the transiion in the experience buffer.
    experience$push(
      s_t = s_t,           # previous state
      s_t1 = s_t1,         # new state
      terminal = terminal, # indicates if the game ended
      action = action,     # stores the performed action
      reward = reward      # the reward obtained by performing the action
    )

    # performs a train step (if selected)
    if (train) 
      train_step(models, experience, env$action_space$n)

    # update state and score
    s_t <- s_t1
    score <- score + reward
  }

  score
}
```

Since `experience` and `models` are modified in place we don't need to return them to persist their values. Note that unlike many other R functions `play_episode` has lot's of side-effects.
